---
title: "Project"
author: "Luis G"
date: "2025-11-10"
output: 
  html_document: 
    highlight: espresso
    theme: readable
---
# Project Plan: Meteorite Landings 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyr)
library(dplyr)
library(stringr)
```


```{r}
meteroite<- read.csv("~/Downloads/Data Science 1/Project/meteorite-landings.csv")

meteroite <- meteroite %>% 
  drop_na(mass) %>% 
  rename(class = recclass)

```

## Source: NASA Open Data ('https://data.nasa.gov/dataset/meteorite-landings')
## This dataset contains detailed records of over 45,000 meteorites that have been found or observed falling to Earth. 

## The dataset includes the following key variables:
- name: Official meteorite name
- id: identification number from NASA
- nametype: “Valid” is a normal meteoroid. “Relict” is a weathered down meteoroid
- recclass: Meteorite classification (e.g., H5, L6, Iron, CM2)
- mass (g): Mass of the meteorite in grams
- fall: “Fell” (observed fall) or “Found” later
- year: The year it fell or was found
- reclat / reclong: Latitude and longitude of discovery site
- GeoLocation: Combined coordinate or location label

## For my project, I'll explore NASA’s Meteorite Landings dataset. The data shows meteorite discoveries from all over the world, including their size, location, and type. I want to find patterns in when and where meteorites are found and what kinds are most common.

## Here are a few questions I want to answer:
- 1. Which meteorite types appear most often?
- 2. Where on Earth have the largest meteorites been found?
- 3. How has the number of meteorite discoveries changed over time?
- 4. Which hemisphere has more recorded meteorites?
- 5. Which countries or regions report the most meteorites?
- 6. Does latitude affect where meteorites land

```{r}
meteroite %>% 
  mutate(
    type = case_when(
      str_starts(class, regex("I", ignore_case = TRUE)) ~ "Iron",
      str_starts(class, regex("P|M", ignore_case = TRUE)) ~ "Stony-Iron",
      str_starts(class, regex("A|U|E|Lun|Mar|D|Hed|How|Euc", ignore_case = TRUE)) ~ "Achondrite",
      str_starts(class, regex("H|L|C|R|E", ignore_case = TRUE)) ~ "Stony",
      TRUE ~ "Other/Unknown"
    ), .before = 3
  )
```

